Score-based tests of measurement invariance: Use in practice

### 2014-07-07 frontiers/IMPSpres-knitr.Rnw
update presentation and lines.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@349 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update presentation.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@350 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Minor changes to IMPS presentation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@351 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%--------------------------------------------------------------------------
\begin{frame}
\frametitle{Measurement Invariance}
  \begin{itemize}
    \item Measurement invariance: Sets of tests/items consistently
      assigning scores across diverse groups of individuals.\\ \ \\
    \item Applications illustrating the importance of measurement invariance:
      \begin{itemize}
         \item SAT for different ethnic groups \cite{atk01}
         \item Intelligence tests' use over time. \cite{wich04}
      \end{itemize}
  \end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Example: Grade 5 vs. Grade 6}
 \begin{figure}
    \label{fig:pathinmodeltheory}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

@@

\node[error] (E1) at (7,6) {E1}
   edge [post] node[auto,swap]{\small{1}}(SCALE1)
   edge [errorsling, bend right=100]node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{11}=26.77$}} (E1.45);
\node[error] (E2) at (7,4.5) {E2}
   edge [post] node[auto,swap]{\small{1}} (SCALE2)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{22}=13.01$}}(E2.45);
\node[error] (E3) at (7,3) {E3}
   edge [post] node[auto,swap]{\small{1}}(SCALE3)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{33}=30.93$}}(E3.45);
\node[error] (E4) at (7,1) {E4}
   edge [post] node[auto,swap]{\small{1}}(SCALE4)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{44}=3.17$}}(E4.45);
\node[error] (E5) at (7,-0.5) {E5}
   edge [post] node[auto,swap]{\small{1}}(SCALE5)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{55}=8.82$}} (E5.45);
\node[error] (E6) at (7,-2) {E6}
   edge [post] node[auto,swap]{\small{1}}(SCALE6)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{66}=22.5$}}(E6.45);
\end{tikzpicture}}
\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Hypotheses}
 Hypothesis of  measurement invariance and alternative:
  \begin{columns}
    \begin{column}{0.48\textwidth}
      \begin{block}{2 Groups}
        \begin{align}
          &H_0: \underset{i\in\left\{1:n\right\}}{\bm{\theta}_i} = 
          \bm{\theta}_0,\\
          &H_1:
           \begin{cases}
               \underset{i \in \left\{v_{1}\right\}}{\bm{\theta}_i} = 
               \bm{\theta}_{v_{1}} \\
               \underset{i \in \left\{v_{2}\right\}}{\bm{\theta}_i} = 
               \bm{\theta}_{v_{2}}, 
           \end{cases}
        \end{align}
       % \begin{center}
       %  where $\left\{v_{1}\right\}+\left\{v_{2}\right\}=\left\{1:n\right\}$
       % \end{center}
      \end{block}
  \end{column}
  \begin{column}{0.48\textwidth}
    \begin{block}{G Groups}
         \begin{align}
          &H_0: \underset{i\in\left\{1:n\right\}}{\bm{\theta}_i} = 
                \bm{\theta}_0,\\
          &H_1: \underset{i \in \left\{v_{g}\right\}}{\bm{\theta}_i} = 
                \bm{\theta}_{v_{g}}, 
         \end{align}
      % \begin{center}
      %  where $\sum_{g=1}^{G}\left\{v_{g}\right\}=\left\{1:n\right\}$
      % \end{center}   
       $V_{g}$ can be categorical, \textit{ordinal} and continous.
    \end{block}
  \end{column}
\end{columns}

\end{frame}
%--------------------------------------------------------------------------
%\begin{frame}
%\frametitle{Problem Scenario}
%\begin{itemize}
%  \item
% \emph{Data:} \\
% The data consist of responses to a gratitude scale from $n=1401$ youth 
% aged 10--19 years, dividing into six age groups: 10-11, 12-13, 14,
% 15, 16, 17-19. 
%  \item
%\emph{Question:} \\
%The researcher is interested in whether the scale
%is measurement invariant across these age groups.  
%\end{itemize}
%
%\end{frame}
%%--------------------------------------------------------------------------
%\begin{frame}
%  \frametitle{Similar Problem Scenarios and Generalization}
%  \begin{itemize}
%    \item Whether model parameter(s) hold(s) for people with different
%      income level, grade level, motivation level, etc.  
%    \item In general, the problem is whether measurement invariance(M.I.)
%      hypothesis holds with respect to an auxiliary ordinal variable. 
%    \end{itemize}
%\end{frame}
%%--------------------------------------------------------------------------
\begin{frame}
\frametitle{Possible Solution}
\begin{itemize}
\item Likelihood Ratio Test (LRT): Require estimates under null
  hypothesis and alternative.
%  \begin{itemize}
%    \item Ignore the nature of ordinal variable, treat it as
%      categorical varaible. 
%    \item Sensitive to sample size. Rejecte the reduced model(s) too often.
%    \item Rather tedious and time consuming to fit and compare all possible
%      models. e.g, for testing one parameter, you need to fit $2^{m-1}$ 
%      models, $m$ is the level of ordinal variable.
%    \end{itemize}
\item Wald Test: Require estimates under alternative. 
\item Lagrange Multiplier Test (Score Test): Require estimates under
  null hypotheis.
\item \textbf{Score Test generalizations w.r.t. general auxiliary
    variable (categorical, ordinal, or continuous)}. 
\end{itemize}
\end{frame}
%  --------------------------------------------------------------------------
\begin{frame}
\frametitle{Outline} 
% Table of contents slide, comment this block out to remove it
\tableofcontents 
% Throughout your presentation, if you choose to use \section{} and 
%\subsection{} commands, these will automatically be printed on this slide 
%as an overview of your presentation
\end{frame}
%  --------------------------------------------------------------------------
\section{Theoretical Background}
\begin{frame}
  \frametitle{Score-based Tests}
  \begin{itemize}
    \item The score-based tests \cite{MerZei13, MerFanZei}
      rely on first derivatives of
      the model's log-likelihood function.
    \item We consider individual terms ({\em scores}) of the gradient.  
      These scores tell us how well a particular parameter describes a 
      particular individual.
  % Equation of gradient vs scores for individuals
  \end{itemize}

  \begin{eqnarray}
  \hat{\bm \theta}=\argmax_{\bm \theta} \ell(\bm \theta; x_{1}, \ldots, x_{n})\\
   s(\hat{{\bm \theta}}; \bm{x}_i) = \frac{\partial}{\partial \bm{\theta}} 
  \log
\text{L}({\bm{x}}_i, {\bm{\theta}}) \big |_{\bm{\theta} = \widehat{\bm{\theta}}}\\
  \displaystyle\sum_{i=1}^{n} s(\hat{\bm{\theta}} ; \bm{x}_i) =
  {\bm 0} 
  \end{eqnarray}

\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Score Aggregation}
  \begin{itemize}
    \item Each individual has a score for each model parameter,
      resulting in $nk$ total scores (where $k$ is the dimension of
      $\bm{\theta}$).
    \item To draw general conclusions about measurement invariance, we
      need to aggregate across scores in specific manners.
    \item When $V$ is ordinal or continuous, we can make use of the
      ordering in $V$ to obtain novel test statistics.
  \end{itemize}
\end{frame}

@@

\vspace{0.2in}
where $\hat{\bm I}$ is an estimate of the information matrix, $\lfloor
nt \rfloor$ is the integer part of $nt$, $t$ is the cumulative proportion 
of $n$, $t \in (1/n, n)$.
  \end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Visual Explanation of the Score-based Test}
  \begin{columns}
    \begin{column}{0.48\textwidth}
      \begin{block}{Score}
        \resizebox{5.0cm}{!}{
        $$\bordermatrix{\text{}& \theta_{1} & \theta_{2} & \ldots 
          &\textcolor{red}{\theta_{k}}\cr
                x_{1}&  1.32  & -4.32 & \ldots & -2.34\cr
                x_{2}&  3.37 & 8.32 & \ldots & 63.45\cr
                x_{3}&  4.12 & 3.48 & \ldots & -3.11\cr
                x_{..}& \vdots & \vdots & \vdots & \vdots\cr
                x_{n-1}&  12.34 & 13.32 & \ldots & 6.89\cr
                x_{n}& -5.78     &   3.50     & \ldots & 7.67} $$}
        \end{block}
     \end{column}
     \begin{column}{0.48\textwidth}
       \begin{block}{Empirical Cumulative Score}
         \resizebox{5.0cm}{!}{
      $$\bordermatrix{\text{}& \theta_{1} & \theta_{2} & \ldots
        &\textcolor{red}{\theta_{k}}\cr
                x_{1}&  1.32  & -4.32 & \ldots & -2.34\cr
                x_{2}&  4.69 & 4.00 & \ldots & 61.11\cr
                x_{3}&  8.81 & 7.48 & \ldots & 58.00\cr
                x_{..}& \vdots & \vdots & \vdots & \vdots\cr
                x_{n-1}&  7.89 & 3.32 & \ldots & 4.91\cr
                x_{n}& 0     &   0     & \ldots & 0} $$}
        \end{block}
      \end{column}
   \end{columns}
\end{frame}
%--------------------------------------------------------------------------

\begin{frame}
  \frametitle{Distributional Theory}
  \begin{itemize}
   \item  Under the hypothesis of measurement invariance, a central limit 
     theorem can be
   used to show that the fluctuation of the above empirical cumulative sum 
   score
   matrix ${\bm B}(\cdot; \hat {\bm \theta})$ converges in distribution to a 
   k-dimensional Brownian bridge ${\bm B}^{0}(\cdot)$.
   \begin{equation}
     \label{eq:converge}
    {\bm B}(\cdot; \hat {\bm \theta}) ~\overset{d}
        {\rightarrow}~ {\bm B}^{0}(\cdot),
   \end{equation}
   \item This result allows us to calculate $p$-values and
   critical values for test statistics under the hypothesis of
   measurement invariance.
  \end{itemize}
 \end{frame}
%--------------------------------------------------------------------------  
 \begin{frame}
   \frametitle{Test Statistics}
\begin{itemize}
  \item Categorical $V$: Maximum Lagrange multiplier test (asymptotically equvialent to LRT).
\item Ordinal $V$: Weighted double maximum test ($\mathit{WDM}_o$);\\
Maximum Lagrange multiplier test ($\max\mathit{LM}_o$).
\item Continuous $V$: Double maximumum test ($\mathit{DM}$)\\
Cram{\'e}r-von Mises type statistic ($\mathit{CvM}$)\\
maximum Lagrange multiplier ($\max\mathit{LM}$)\\
\end{itemize}
\end{frame}

@@

%-------------------------------------------------------------------------   
\section{Simulation}
\begin{frame}
  \frametitle{Study Motivation}
  \begin{block}{Practical Considerations}
    \begin{itemize}
      \item Model misspecification is inevitable in practice.  
      \item Measurement invariance is often studied with respect to an
        \textbf{ordinal} variable.
      \item Availability/Convenience of software implementation.
    \end{itemize}
  \end{block}
\end{frame}
%-------------------------------------------------------------------------   
\begin{frame}
  \frametitle{Simulation Motivation}
  \begin{block}{Simulation:}
    What are the ordinal tests' power under model misspecification? 
    \begin{itemize}
      \item The estimated model excludes one parameter, which is \emph
        {not} the problematic parameter.
      \item The estimated model excludes one parameter, which \emph{is}
        the problematic parameter. 
    \end{itemize}
  \end{block}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Simulation Setup}
 \begin{itemize}
   \item Three sample sizes ($n=120,\textcolor{red}{480},960$), 
   \item Three numbers of ordinal levels ($m=4,8,12$), 
   \item Violation magnitude: $d$ times the parameters' asymptotic 
     standard errors (scaled by $\sqrt n$), with $d=0,0.25,0.5,...,4$ 
   \item Ordinal Violation: began at level $1+m/2$ of the auxiliary 
     variable $V$.
 \end{itemize}
   Statistics $\mathit{WDM}_o$, $\max \mathit{LM}_o$, $\mathit{LM}_{uo}$ were 
   examined.
\end{frame}
%--------------------------------------------------------------------------

@@

\node[error] (E1) at (7,6) {E1}
   edge [post] node[auto,swap]{\small{1}}(SCALE1)
   edge [errorsling, bend right=100]node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{11}=26.77$}} (E1.45);
\node[error] (E2) at (7,4.5) {E2}
   edge [post] node[auto,swap]{\small{1}} (SCALE2)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{22}=13.01$}}(E2.45);
\node[error] (E3) at (7,3) {E3}
   edge [post] node[auto,swap]{\small{1}}(SCALE3)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{33}=30.93$}}(E3.45);
\node[error] (E4) at (7,1) {E4}
   edge [post] node[auto,swap]{\small{1}}(SCALE4)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{44}=3.17$}}(E4.45);
\node[error] (E5) at (7,-0.5) {E5}
   edge [post] node[auto,swap]{\small{1}}(SCALE5)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{55}=8.82$}} (E5.45);
\node[error] (E6) at (7,-2) {E6}
   edge [post] node[auto,swap]{\small{1}}(SCALE6)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{66}=22.5$}}(E6.45);
\end{tikzpicture}}
\end{figure}
\end{frame}
%--------------------------------------------------------------------------

@@

\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{$\lambda_{11}$ violates measurement invariance.  Tested
    parameters are $\phi_{12}$, \{$\psi_{11}, \ldots, \psi_{66}$\}, and
    \{$\mu_{11}, \ldots, \mu_{66}$\}.}}
\label{fig:sim22res}
<<sim22res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra+loading" & 
          pars %in% c("All Errors", "Covariance", "All Intercepts") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(2, 6, 7)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

\end{figure}
\end{frame}
%--------------------------------------------------------------------------

>>>>

\begin{frame}
\frametitle{Measurement Invariance}
  \begin{itemize}
    \item Measurement invariance: Sets of tests/items consistently
      assigning scores across diverse groups of individuals.\\ \ \\
    \item Applications illustrating the importance of measurement invariance:
      \begin{itemize}
         \item SAT for different ethnic groups \cite{atk01}
         \item Intelligence tests' use over time. \cite{wich04}
      \end{itemize}
  \end{itemize}
\end{frame}

% --------------------------------------------------------------------------

\begin{frame}
  \frametitle{Example: Grade 5 vs. Grade 6}
    \begin{figure}
      \label{fig:pathinmodeltheory}
         \resizebox{8.0cm}{!}{
           \begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
inner sep=0pt, minimum width=20mm, minimum height=10mm},
manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
inner sep=0pt,minimum width=20mm, minimum height=10mm},
error/.style={circle,draw=black!50,fill=white!20,thick,
inner sep=0pt, minimum width=10mm},
post/.style={->,shorten >=1pt,>=stealth',semithick},
postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
auto,distance=20mm,semithick},
errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
auto,distance=10mm,semithick}]

@@

\node[error] (E1) at (7,6) {E1}
   edge [post] node[auto,swap]{\small{1}}(SCALE1)
   edge [errorsling, bend right=100]node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{11}=26.77$}} (E1.45);
\node[error] (E2) at (7,4.5) {E2}
   edge [post] node[auto,swap]{\small{1}} (SCALE2)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{22}=13.01$}}(E2.45);
\node[error] (E3) at (7,3) {E3}
   edge [post] node[auto,swap]{\small{1}}(SCALE3)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{33}=30.93$}}(E3.45);
\node[error] (E4) at (7,1) {E4}
   edge [post] node[auto,swap]{\small{1}}(SCALE4)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{44}=3.17$}}(E4.45);
\node[error] (E5) at (7,-0.5) {E5}
   edge [post] node[auto,swap]{\small{1}}(SCALE5)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{55}=8.82$}} (E5.45);
\node[error] (E6) at (7,-2) {E6}
   edge [post] node[auto,swap]{\small{1}}(SCALE6)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{66}=22.5$}}(E6.45);
\end{tikzpicture}}
\end{figure}
\end{frame}

% --------------------------------------------------------------------------

\begin{frame}
  \frametitle{Hypotheses}
 Hypothesis of  measurement invariance and alternative:
  \begin{columns}
    \begin{column}{0.48\textwidth}
      \begin{block}{2 Groups}
        \begin{align}
          &H_0: \underset{i\in\left\{1:n\right\}}{\bm{\theta}_i} = 
          \bm{\theta}_0,\\
          &H_1:
           \begin{cases}
               \underset{i \in \left\{v_{1}\right\}}{\bm{\theta}_i} = 
               \bm{\theta}_{v_{1}} \\
               \underset{i \in \left\{v_{2}\right\}}{\bm{\theta}_i} = 
               \bm{\theta}_{v_{2}}, 
           \end{cases}
        \end{align}
       % \begin{center}
       %  where $\left\{v_{1}\right\}+\left\{v_{2}\right\}=\left\{1:n\right\}$
       % \end{center}
      \end{block}
  \end{column}
  \begin{column}{0.48\textwidth}
    \begin{block}{G Groups}
         \begin{align}
          &H_0: \underset{i\in\left\{1:n\right\}}{\bm{\theta}_i} = 
                \bm{\theta}_0,\\
          &H_1: \underset{i \in \left\{v_{g}\right\}}{\bm{\theta}_i} = 
                \bm{\theta}_{v_{g}}, 
         \end{align}
      % \begin{center}
      %  where $\sum_{g=1}^{G}\left\{v_{g}\right\}=\left\{1:n\right\}$
      % \end{center}   
       $V_{g}$ can be categorical, \textit{ordinal} and continous.
    \end{block}
  \end{column}
\end{columns}

\end{frame}

% --------------------------------------------------------------------------

% \begin{frame}
%\frametitle{Problem Scenario}
%\begin{itemize}
%  \item
% \emph{Data:} \\
% The data consist of responses to a gratitude scale from $n=1401$ youth 
% aged 10--19 years, dividing into six age groups: 10-11, 12-13, 14,
% 15, 16, 17-19. 
%  \item
%\emph{Question:} \\
%The researcher is interested in whether the scale
%is measurement invariant across these age groups.  
%\end{itemize}
%
%\end{frame}
%%--------------------------------------------------------------------------
%\begin{frame}
%  \frametitle{Similar Problem Scenarios and Generalization}
%  \begin{itemize}
%    \item Whether model parameter(s) hold(s) for people with different
%      income level, grade level, motivation level, etc.  
%    \item In general, the problem is whether measurement invariance(M.I.)
%      hypothesis holds with respect to an auxiliary ordinal variable. 
%    \end{itemize}
%\end{frame}

%%--------------------------------------------------------------------------

\begin{frame}
\frametitle{Possible Solution}
\begin{itemize}
\item Likelihood Ratio Test (LRT): Require estimates under null
  hypothesis and alternative.
%  \begin{itemize}
%    \item Ignore the nature of ordinal variable, treat it as
%      categorical varaible. 
%    \item Sensitive to sample size. Rejecte the reduced model(s) too often.
%    \item Rather tedious and time consuming to fit and compare all possible
%      models. e.g, for testing one parameter, you need to fit $2^{m-1}$ 
%      models, $m$ is the level of ordinal variable.
%    \end{itemize}
\item Wald Test: Require estimates under alternative. 
\item Lagrange Multiplier Test (Score Test): Require estimates under
  null hypotheis.
\item \textbf{Score Test generalizations w.r.t. general auxiliary
    variable (categorical, ordinal, or continuous)}. 
\end{itemize}
\end{frame}

% --------------------------------------------------------------------------

\begin{frame}
\frametitle{Outline} 
% Table of contents slide, comment this block out to remove it
\tableofcontents 
% Throughout your presentation, if you choose to use \section{} and 
%\subsection{} commands, these will automatically be printed on this slide 
%as an overview of your presentation
\end{frame}

% --------------------------------------------------------------------------

\section{Theoretical Background}
\begin{frame}
  \frametitle{Score-based Tests}
 \begin{itemize}
    \item The score-based tests \cite{MerZei13, MerFanZei}
      rely on first derivatives of
      the model's log-likelihood function.
    \item We consider individual terms ({\em scores}) of the gradient.  
      These scores tell us how well a particular parameter describes a 
      particular individual.
  % Equation of gradient vs scores for individuals
  \end{itemize}

\scriptsize{   
  \begin{eqnarray}
    \label{eq:scoreml}
    s(\hat{{\bm \theta}}; \bm{x}_i) = \frac{\partial}{\partial \bm{\theta}} 
    \log\text{L}({\bm{x}}_i, {\bm{\theta}}) \big |_{\bm{\theta} =
    \widehat{\bm{\theta}}}, \text{where}\\[1em]
    \label{eq:ml}
    \hat{\bm \theta}=\argmax_{\bm \theta} \log\text{L}(\bm \theta; x_{1}, 
    \ldots, x_{n})\\
    \label{eq:scoresumml}
   \displaystyle\sum_{i=1}^{n} s(\hat{\bm{\theta}} ; \bm{x}_i) = {\bm 0}
  \end{eqnarray}
}

\end{frame}

% --------------------------------------------------------------------------

\begin{frame}
  \frametitle{Score Aggregation}
  \begin{itemize}
    \item Each individual has a score for each model parameter,
      resulting in $nk$ total scores (where $k$ is the dimension of
      $\bm{\theta}$).
    \item To draw general conclusions about measurement invariance, we
      need to aggregate across scores in specific manners.
    \item When $V$ is ordinal or continuous, we can make use of the
      ordering in $V$ to obtain novel test statistics.
  \end{itemize}
\end{frame}

@@

\vspace{0.2in}
where $\hat{\bm I}$ is an estimate of the information matrix, $\lfloor
nt \rfloor$ is the integer part of $nt$, $t$ is the cumulative proportion 
of $n$, $t \in (1/n, n)$.
  \end{frame}

% --------------------------------------------------------------------------

\begin{frame}
  \frametitle{Visual Explanation of the Score-based Test}
  \begin{columns}
    \begin{column}{0.48\textwidth}
      \begin{block}{Score}
        \resizebox{5.0cm}{!}{
        $$\bordermatrix{\text{}& \theta_{1} & \theta_{2} & \ldots 
          &\textcolor{red}{\theta_{k}}\cr
                x_{1}&  1.32  & -4.32 & \ldots & -2.34\cr
                x_{2}&  3.37 & 8.32 & \ldots & 63.45\cr
                x_{3}&  4.12 & 3.48 & \ldots & -3.11\cr
                x_{..}& \vdots & \vdots & \vdots & \vdots\cr
                x_{n-1}&  12.34 & 13.32 & \ldots & 6.89\cr
                x_{n}& -5.78     &   3.50     & \ldots & 7.67} $$}
        \end{block}
     \end{column}
     \begin{column}{0.48\textwidth}
       \begin{block}{Empirical Cumulative Score}
         \resizebox{5.0cm}{!}{
      $$\bordermatrix{\text{}& \theta_{1} & \theta_{2} & \ldots
        &\textcolor{red}{\theta_{k}}\cr
                x_{1}&  1.32  & -4.32 & \ldots & -2.34\cr
                x_{2}&  4.69 & 4.00 & \ldots & 61.11\cr
                x_{3}&  8.81 & 7.48 & \ldots & 58.00\cr
                x_{..}& \vdots & \vdots & \vdots & \vdots\cr
                x_{n-1}&  7.89 & 3.32 & \ldots & 4.91\cr
                x_{n}& 0     &   0     & \ldots & 0} $$}
        \end{block}
      \end{column}
   \end{columns}
\end{frame}

% --------------------------------------------------------------------------

\begin{frame}
  \frametitle{Distributional Theory}
  \begin{itemize}
   \item  Under the hypothesis of measurement invariance, a central limit 
     theorem can be
   used to show that the fluctuation of the above empirical cumulative sum 
   score
   matrix ${\bm B}(\cdot; \hat {\bm \theta})$ converges in distribution to a 
   $k$-dimensional Brownian bridge ${\bm B}^{0}(\cdot)$.
   \begin{equation}
     \label{eq:converge}
    {\bm B}(\cdot; \hat {\bm \theta}) ~\overset{d}
        {\rightarrow}~ {\bm B}^{0}(\cdot),
   \end{equation}
   \item This result allows us to calculate $p$-values and
   critical values for test statistics under the hypothesis of
   measurement invariance.
  \end{itemize}
 \end{frame}

% --------------------------------------------------------------------------

\begin{frame}
   \frametitle{Test Statistics}
\begin{itemize}
  \item Categorical $V$: Maximum Lagrange multiplier test (asymptotically equvialent to LRT).
\item Ordinal $V$: Weighted double maximum test ($\mathit{WDM}_o$);\\
Maximum Lagrange multiplier test ($\max\mathit{LM}_o$).
\item Continuous $V$: Double maximumum test ($\mathit{DM}$)\\
Cram{\'e}r-von Mises type statistic ($\mathit{CvM}$)\\
maximum Lagrange multiplier ($\max\mathit{LM}$)\\
\end{itemize}
\end{frame}

@@

%-------------------------------------------------------------------------

\section{Simulation}
\begin{frame}
  \frametitle{Study Motivation}
  \begin{block}{Practical Considerations}
    \begin{itemize}
      \item Model misspecification is inevitable in practice.  
      \item Measurement invariance is often studied with respect to an
        \textbf{ordinal} variable.
      \item Availability/Convenience of software implementation.
    \end{itemize}
  \end{block}
\end{frame}

% -------------------------------------------------------------------------

\begin{frame}
  \frametitle{Simulation Motivation}
  \begin{block}{Simulation:}
    What are the ordinal tests' power under model misspecification? 
    \begin{itemize}
      \item The estimated model excludes one parameter, which 
        {\em is not} the problematic parameter.
      \item The estimated model excludes one parameter, which \emph{is}
        the problematic parameter. 
    \end{itemize}
  \end{block}
\end{frame}

% --------------------------------------------------------------------------

\begin{frame}
  \frametitle{Simulation Setup}
 \begin{itemize}
   \item Three sample sizes ($n=120,\textcolor{red}{480},960$), 
   \item Three numbers of ordinal levels ($m=4,8,12$), 
   \item Violation magnitude: $d$ times the parameters' asymptotic 
     standard errors (scaled by $\sqrt n$), with $d=0,0.25,0.5,...,4$ 
   \item Ordinal Violation: began at level $1+m/2$ of the auxiliary 
     variable $V$.
 \end{itemize}
   Statistics $\mathit{WDM}_o$, $\max \mathit{LM}_o$, $\mathit{LM}_{uo}$ were 
   examined.
\end{frame}

% --------------------------------------------------------------------------

@@

\node[error] (E1) at (7,6) {E1}
   edge [post] node[auto,swap]{\small{1}}(SCALE1)
   edge [errorsling, bend right=100]node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{11}=26.77$}} (E1.45);
\node[error] (E2) at (7,4.5) {E2}
   edge [post] node[auto,swap]{\small{1}} (SCALE2)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{22}=13.01$}}(E2.45);
\node[error] (E3) at (7,3) {E3}
   edge [post] node[auto,swap]{\small{1}}(SCALE3)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{33}=30.93$}}(E3.45);
\node[error] (E4) at (7,1) {E4}
   edge [post] node[auto,swap]{\small{1}}(SCALE4)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{44}=3.17$}}(E4.45);
\node[error] (E5) at (7,-0.5) {E5}
   edge [post] node[auto,swap]{\small{1}}(SCALE5)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{55}=8.82$}} (E5.45);
\node[error] (E6) at (7,-2) {E6}
   edge [post] node[auto,swap]{\small{1}}(SCALE6)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{66}=22.5$}}(E6.45);
\end{tikzpicture}}
\end{figure}
\end{frame}

% --------------------------------------------------------------------------

@@

\end{figure}
\end{frame}

% --------------------------------------------------------------------------

\begin{frame}
\begin{figure}
\tiny{\caption{$\lambda_{11}$ violates measurement invariance.  Tested
    parameters are $\phi_{12}$, \{$\psi_{11}, \ldots, \psi_{66}$\}, and
    \{$\mu_{11}, \ldots, \mu_{66}$\}.}}
\label{fig:sim22res}
<<sim22res, fig.height=5, fig.width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra+loading" & 
          pars %in% c("All Errors", "Covariance", "All Intercepts") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(2, 6, 7)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

\end{figure}
\end{frame}

% --------------------------------------------------------------------------
